Tests of measurement invariance via stochastic processes: {S}ome properties

### 2013-07-17 frontiers/frontiers.Rnw
update frontiers to show code in Rnw file
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@216 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
Minor revisions to frontiers paper
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@217 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
update frontier with lavaan example
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@218 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\SweaveOpts{engine = R, eps = FALSE, echo = FALSE, results = hide}

@@

\section{Tutorial}
\section{Simulation 1}

>>>>

\SweaveOpts{engine = R, eps = FALSE, echo = FALSE, pdf=TRUE}

@@

\SweaveOpts{engine = R, eps = FALSE, echo = TRUE, pdf=TRUE}
\section{Tutorial}
In general, a set of scales is defined to be measurement invariant with respect to an auxiliary variable $V$ if:

\begin{equation}
\label{equ:invariance}
f(x_{i}|t_{i},v_{i},...)=f(x_{i}|t_{i},...),
\end{equation}

where $x_{i}$ is the data vector for individual $i$, $t_{i}$ is the latent variable for variable $i$ that the scales purport to measure, and $f$ is the model's distributional form \citep{}. If the cut point of $v$ is known in advance, i.e. gender, nested multiple group models\citep{} coupled with likelihood ratio test are most commonly applied. 
Suppose there are two groups A and B. The null hypothesis of measurement invariance is

\begin{equation}
\label{equ:null}
H_{0}: \theta_{A}=\theta_{B},
\end{equation}

The alternative hypothesis is

\begin{equation}
\label{equ:alternative}
H_{1}:\theta_{i}=
\begin{cases}
\theta^{(A)} & v_{i}\leq v\\
\theta^{(B)} & v_{i}\geq v
\end{cases}
\end{equation}

The parameter estimate $\hat{\theta}^{(A)}$ , $\hat{\theta}^{(B)}$ ,$\hat{\theta}$ (group A and group B) can be obtained by Maximum Likelihood(ML) from observations $i=1,...m(v_{i}\leq v),i=m+1,...n(v_{i}\geq v), i=1,...,n$,respectively. The LR test statistic for the given threshold $v$ is then 
\begin{equation}
\label{equ:LRT}
LR(v)=-2[\ell(\hat{\theta};x_{1},,,,x_{n})-{\ell(\hat{\theta}^
{(A)};x_{1},...,x_{m})+\ell(\hat{\theta}^{(B)};x_{m+1},...,x_{n})
}],
\end{equation}

has an asymptotic $\chi^{2}$ distribution with degrees of freedom equal to the number difference of parameters in $\theta$.

To be more specific, in the following section, we will introduce how to implement multi group analysis in lavaan package. For more details, see\citep{}
\subsection{Multiple Group Analysis}
First of all, we need to read in data to R.  
<<read data>>=
# Reading in the data
## (need data here)Data <- read.table("example.dat")
## names(Data) <- c("y1","y2","y3","y4","y5","y6","g")
@
Second, the model need to be specified using lavaan package. 
<<specify model>>=
library(lavaan)
model <- 'f1=~NA*y1+y2+y3
          f2=~NA*y4+y5+y6
          f1 ~~1*f1
          f2 ~~1*f2'
@
Third, we want to test which parameter(s) violate measurement invariance. In lavaan package, loading, manifest variables' intercept and factor means could be tested together or specifically. For more details, please see \citep{}. Since in most cases researchers are interested in the factor loading invariance, we will use that as the example in this tutorial.
To implement LRT as ~\ref{equ:likelihood}, we need to get estimation for \hat{\theta}
<<theta estimate>>=
# estimate theta 
## theta<-cfa(model,data=Data,group="g",group.equal="loadings")
@
Then we can get estimate for group A and group B. 
<<theta A and theta B estimate>>=
# estimate theta A,B 
## thetaAB<-cfa(model,data=Data,group="g")
@
In the end, we can get the LRT results by calling the measurement invariance function in lavaan. 
<<LRT>>=
# LRT
## measurementInvariance(model,data=Data,group="g",strict=FALSE)
@
\subsection{Extensions for Unknown Group}

\SweaveOpts{engine = R, eps = FALSE, echo = FALSE, pdf=TRUE}
\section{Simulation 1}
